Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-06-04 Manuscript/Figures/figure_legends.Rmd

**Figure 1. Model Construction.** An illustration of the proposed model construction given cell line specific Hi-C data. Above is an example of a subsection of the contact matrix for chromosome 14 (at megabase scale) on GM12878 [@durand2016juicer]. The contact matrix is a visualization of the interconnectivity of the linear genome at 100 kb resolution. The boundary coordinates of TADs were identified using the Arrowhead algorithm. To account for the uncertainty in TAD boundary detection, each boundary point was flanked on both sides by resolution-specific genomic regions (bins). The response vector $Y$, used for classification, was determined by whether a genomic bin overlapped with a flanked boundary region ($Y=1$) or not ($Y=0$). The spatial relationship between positional coordinates of various functional genomic annotations, obtained from the UCSC Genome Browser Database, together with the genomic bins were then used to define the feature space of the models represented by the data matrix $X = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations.

@@

**Figure 3. Ensemble Framework.** A diagram of the cell line specific ensemble framework detailing the combination of inputs used in the model building pipeline. The minoirty and majority classes were represented by the genomic bins that overlapped with flanked TAD boundaries or did not, respectively. Functional genomic annotations obtained from the UCSC Genome Browser Database composed the feature space ($X = \{X_1, X_2, ..., X_p\}$). A feature engineering step was included to isolate the predictor-type specific spatial relationship between annotations and genomic bins in increasing granularity, denoted by the red gradient. The data was split into a 7:3 training to testing set ratio. Data re-sampling was then applied to the training set. Lastly, a random forest classification algorithm was implemeted and validated on the testing data.

>>>>

**Figure 1. Model Construction.** An illustration of the proposed model construction given cell line specific Hi-C data. Above is an example of a subsection of the contact matrix for chromosome 14 (at megabase scale) on GM12878 [@durand2016juicer]. The contact matrix is a visualization of the interconnectivity of the linear genome at 100 kb resolution. The boundary coordinates of TADs were identified using the Arrowhead algorithm. To account for the uncertainty in TAD boundary detection, each boundary point was flanked on both sides by resolution-specific genomic regions (bins). The response vector $Y$, used for classification, was determined by whether a genomic bin overlapped with a flanked boundary region ($Y=1$) or not ($Y=0$). The spatial relationship between positional coordinates of various functional genomic annotations, obtained from the UCSC Genome Browser Database, together with the genomic bins were then used to define the feature space of the models represented by the data matrix $\textbf{X} = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations.

@@

**Figure 3. Ensemble Framework.** A diagram of the cell line specific ensemble framework detailing the combination of inputs used in the model building pipeline. The minoirty and majority classes were represented by the genomic bins that overlapped with flanked TAD boundaries or did not, respectively. Functional genomic annotations obtained from the UCSC Genome Browser Database composed the feature space ($\textbf{X} = \{X_1, X_2, ..., X_p\}$). A feature engineering step was included to isolate the predictor-type specific spatial relationship between annotations and genomic bins in increasing granularity, denoted by the red gradient. The data was split into a 7:3 training to testing set ratio. Data re-sampling was then applied to the training set. Lastly, a random forest classification algorithm was implemeted and validated on the testing data.

### 2019-06-04 Manuscript/manuscript.Rmd

---
title: Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance
  Settings
csl: styles.ref/genomebiology.csl
output:
  word_document:
    reference_docx: styles.doc/NIH_grant_style.docx
  pdf_document: default
bibliography: References/references.bib
---

@@

@@

@@

To establish the feature space for all downstream models, we constructed 3 different predictor-types. These predictor-type specific annotations represented the data matrix, $\textbf{X} = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations [?? $p$ is actually the total number of functional annotations, $X_i$ is the functional annotation itself ??]. Overlap counts was the least granular [?? Maybe define granular ??] predictor-type considered, simply defined by the number of annotation-specific regions located within each genomic bin (Figure 2A). Overlap percents was a more refined [?? what is refined in this context ??] predictor-type, defined as the total percent overlap between each annotation-specific region and genomic bins (Figure 2B). Lastly, distances, the most granular predictor-type, described the spatial relationship between the center of each genomic bin to the center of the nearest annotation-specific region (Figure 2C). A $log_{2}}$ transformation was used as a normalization technique in order to account for the skewness present in the distance densities (Supplementary Figures 1-2). [?? Why would skewness matter, why not use raw data ??]

@@

>>>>

---
title: Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance
  Settings
csl: styles.ref/genomebiology.csl
output:
  pdf_document: default
  word_document:
    reference_docx: styles.doc/NIH_grant_style.docx
bibliography: References/references.bib
---

@@

![**Model Construction.** An illustration of the proposed model construction given cell line specific Hi-C data. Above is an example of a subsection of the contact matrix for chromosome 14 (at megabase scale) on GM12878 [@durand2016juicer]. The contact matrix is a visualization of the interconnectivity of the linear genome at 100 kb resolution. The boundary coordinates of TADs were identified using the Arrowhead algorithm. To account for the uncertainty in TAD boundary detection, each boundary point was flanked on both sides by resolution-specific genomic regions (bins). The response vector $Y$, used for classification, was determined by whether a genomic bin overlapped with a flanked boundary region ($Y=1$) or not ($Y=0$). The spatial relationship between positional coordinates of various functional genomic annotations, obtained from the UCSC Genome Browser Database, together with the genomic bins were then used to define the feature space of the models represented by the data matrix $\textbf{X} = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations.](C:/Users/stili/Documents/TAD_predictive_modeling/Manuscript/Figures/figure1_model_construction_V3.png)

@@

**Table 1. Domain Data Summaries and Imbalance Ratios.** For GM12878, average number of genomic bins, TADs called, and the corresponding boundaries and imbalance ratios, across each of the autosomal chromosomes (omitting chromosome 9), stratified by resolution. The resolution of the Hi-C contact matrix dictated the width of each genomic bin, with lower resolutions corresponding to wider genomic bins.

| Resolution (Bin Width) | Median Number of Genomic Bins (IQR) | Median Number of TADs (IQR) | Median Number of TAD Boundaries (IQR) | Median Imbalance Ratio (IQR) |
|------------------------|-------------------------------------|-----------------------------|---------------------------------------|------------------------------|
| 10 kb                  | 13055 (8961)                        | 380 (203)                   | 1338 (832)                            | 0.11 (0.02)                  |
| 25 kb                  | 5226 (3586)                         | 220 (155)                   | 737 (521)                             | 0.17 (0.02)                  |
| 50 kb                  | 2616 (1794)                         | 110 (86)                    | 367 (294)                             | 0.17 (0.02)                  |
| 100 kb                 | 1310 (897)                          | 53 (37)                     | 169 (136)                             | 0.15 (0.02)                  |

@@

To establish the feature space for all downstream models, we constructed 3 different predictor-types. These predictor-type specific annotations represented the data matrix, $\textbf{X} = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations [?? $p$ is actually the total number of functional annotations, $X_i$ is the functional annotation itself ??]. Overlap counts was the least granular [?? Maybe define granular ??] predictor-type considered, simply defined by the number of annotation-specific regions located within each genomic bin (Figure 2A). Overlap percents was a more refined [?? what is refined in this context ??] predictor-type, defined as the total percent overlap between each annotation-specific region and genomic bins (Figure 2B). Lastly, distances, the most granular predictor-type, described the spatial relationship between the center of each genomic bin to the center of the nearest annotation-specific region (Figure 2C). A $log_{2}$ transformation was used as a normalization technique in order to account for the skewness present in the distance densities (Supplementary Figures 1-2). [?? Why would skewness matter, why not use raw data ??]

<!--![Caption for the picture.](/path/to/image.png)-->

@@

<!--![Caption for the picture.](/path/to/image.png)-->
